---
editor_options:
  chunk_output_type: console
---

# Temporal sampling

In this script, we examine temporal differences in the reporting of observations for the states of Tamil Nadu and Kerala. We apply an additional filter in this script, where we choose complete checklists only. 


The analyses we carry out include: a) reporting of checklists for every 5-year interval
b) are there significant differences between pre2000 and other time periods

## Prepare libraries
```{r setup_sup_02}
# load libraries
library(data.table)
library(readxl)
library(magrittr)
library(stringr)
library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)
library(ggthemes)
library(scico)
library(extrafont)
library(sf)
library(rnaturalearth)
library(auk)

# round any function
round_25 <- function(x, accuracy = 25000) {
  round(x / accuracy) * accuracy
}

# set file paths for auk functions
# To use these two datasets, please download the latest versions from https://ebird.org/data/download and set the file path accordingly. Since these two datasets are extremely large, we have not uploaded the same to GitHub.
# In this study, the version of data loaded corresponds to July 2022.

f_in_ebd <- file.path("data/ebd_IN_relJul-2022.txt")
f_in_sampling <- file.path("data/ebd_sampling_relJul-2022.txt")
```

## Filter data using auk package

Here, we set broad spatial filters to subset all complete checklsits for the states of Kerala and Tamil Nadu for the above list of species.
```{r prep_ebd_filters}
# run filters using auk packages
ebd_filters <- auk_ebd(f_in_ebd, f_in_sampling) %>%
  auk_country(country = "IN") %>%
  auk_state(c("IN-KL", "IN-TN")) %>%
  auk_complete()

# check filters
ebd_filters
```

Below code need not be run if it has been filtered once already and the above path leads to the right dataset. NB: This is a computation heavy process, run with caution.

```{r output_loc}
# specify output location and perform filter
f_out_ebd <- "data/01_ebird-filtered-KL-TN.txt"
f_out_sampling <- "data/01_ebird-filtered-sampling-KL-TN.txt"
```

```{r filter_data}
ebd_filtered <- auk_filter(ebd_filters,
  file = f_out_ebd,
  file_sampling = f_out_sampling, overwrite = TRUE
)
```

## Process filtered data

The data has been filtered above using the auk functions. We will now work with the filtered checklist observations.  

```{r read_data}
# read in the data
# note: running this takes a very long time and I would suggest saving an .RData file after applying the basic filters
ebd <- read_ebd(f_out_ebd)
gc()

# Convert all presences marked 'X' as '1'
ebd <- ebd %>%
  mutate(observation_count = ifelse(observation_count == "X",
    "1", observation_count))

# Convert observation count to numeric
ebd$observation_count <- as.numeric(ebd$observation_count)

# choose columns of interest
columnsOfInterest <- c(
  "checklist_id", "scientific_name", "common_name",
  "observation_count", "locality", "locality_id",
  "locality_type", "latitude", "longitude",
  "observation_date", "time_observations_started",
  "observer_id", "sampling_event_identifier",
  "protocol_type", "duration_minutes",
  "effort_distance_km", "effort_area_ha",
  "number_observers", "species_observed","state",
  "reviewed"
)
ebd <- ebd %>%
  select(matches(columnsOfInterest))
```

## Include only species with scientific names
```{r}
# remove species that are poorly reported (only filter being applied at this stage)
# for example, accounts such as 'Turdidae sp.', are removed and only the latest
# list of scientific names are taken into consideration

trait_dat <- read.csv("data/2022-SoIB-habitat-data.csv")
speciesOfInterest <- data.frame(trait_dat$eBird.Scientific.Name.2021)
names(speciesOfInterest) <- "scientific_name"

# filter for species list above
ebd <- ebd[ebd$scientific_name %in% speciesOfInterest$scientific_name, ]
```

## Subset data by time period

We will divide data into five-year intervals: 2000-2005, 2006-2010, 2011-2015, 2016-2020 and all data before 2000 as a single category. 

```{r}
# separate observation_date as year, month, day
ebd <- ebd %>%
  separate(observation_date, c("year","month","day"))

# subset data by time period
ebd <- ebd %>%
  mutate(timePeriod = case_when(
    year >= 2000 & year < 2005 ~ "2000-2005",
    year >= 2005 & year < 2010 ~ "2005-2010",
    year >= 2010 & year < 2015 ~ "2010-2015",
    year >= 2015 & year < 2020 ~ "2015-2020",
    year >= 2020 ~ "post2020",
    year < 2000 ~ "pre2000"
  ))
```

## Figure: Overall number of checklists across time-periods per state
```{r}
trial <- ebd %>%
  group_by(sampling_event_identifier) %>%
  mutate()



fig_totChk_timePeriod <- ggplot(ebd, aes(x="timePeriod", y=PXt, fill=Habitat.type)) +  geom_boxplot(alpha=0.7) +  
  scale_fill_scico_d(palette = "roma",
                     labels=c("Forest", "Grassland", "Generalist", "Wooded Habitat")) +   theme_bw() +
  geom_point(aes(fill = Habitat.type), size = 4, shape = 21, alpha = 0.3,position = position_jitterdodge()) +
    labs(x="", 
       y="Persistence probabilty\n") +
  scale_x_discrete(labels = 'Modern') +
    theme(axis.title = element_text(family = "Century Gothic",
      size = 14, face = "bold"),
        axis.text = element_text(family="Century Gothic",size = 14),
      axis.text.x = element_blank(),
      legend.title = element_text(family="Century Gothic",
                                    size = 14, face = "bold"),
        legend.key.size = unit(1,"cm"),
        legend.text = element_text(family="Century Gothic",size = 14))



# facet by State



# visualization of persistence probabilities at the landscape level
specDatTrait$Habitat.type <- factor(specDatTrait$Habitat.type, levels=c("Forest", "Grassland", "Generalist", "Wooded Habitat"))
specDatTrait$PXt <- as.numeric(specDatTrait$PXt)






```





```{r read_data}
# read in shapefile of the study area to subset by bounding box

# first, we load the India states boundary from GADM
india_states <- st_read("data/spatial/shapefiles/IND_adm1.shp")

# load tamil nadu and kerala separately
tamil_nadu <- india_states[india_states$NAME_1=="Tamil Nadu",]
box_tn <- st_bbox(tamil_nadu)

kerala <- india_states[india_states$NAME_1=="Kerala",]
box_ker <- st_bbox(kerala)  

# tamil nadu
tn_ebd <- ebd %>%
  filter(between(longitude, box_tn["xmin"], box_tn["xmax"]) &
  between(latitude, box_tn["ymin"], box_tn["ymax"]))

# tamil nadu
ker_ebd <- ebd %>%
  filter(between(longitude, box_ker["xmin"], box_ker["xmax"]) &
  between(latitude, box_ker["ymin"], box_ker["ymax"]))
```



```{r}



```








